\large{\bf Large effects and the infinitesimal model }

### 2023-07-07 writeup.tex

Unfortunately, once selection is included, \eqref{eqn:MSB_CF} is less amenable to the sort of simple arguments above.  We will content ourselves with presenting two examples in the case of stabilizing selection, one already familiar to enthusiasts of quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.

\subsubsection{Gaussian Solution}
    \label{sec:stabilizing_gaussian}

First, consider the case when $1+ \mu(x) = e^{a x^{2}}$.   In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
\[
	\phi(\theta) = e^{i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}}.
\]
Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]}$  using the Gaussian probability density function to conclude that
if $R$ is Gaussian with mean zero and variance $\eta^2$, then equation~\eqref{eqn:MSB_CF} becomes
\[
	e^{i m \theta - \frac{1}{2} \left(\frac{\sigma^{2}}{2} + \eta^{2}\right)\theta^{2}}
	= e^{i \frac{m}{1-2 a \sigma^{2}}\theta 
		- \frac{1}{2} \frac{\sigma^{2}}{1-2 a \sigma^{2}}\theta^{2}},
\]
which, equating real and complex components, has a solution provided $m = 0$ and
\[
	\sigma^{2} = \frac{\sqrt{16 a^{2} \eta^{4} + 24 a \eta^{2}+1} - 4a \eta^{2} -1}{4a} 
	=  2\eta^{2}(1-4a\eta^{2}) + O(a^{2}).
\]

Note that we do not require very weak selection to guarantee a Gaussian steady state, only that $a < \frac{1}{2\sigma^{2}}$.  However, this does not imply that the transient distribution is also Gaussian, as is sometimes assumed \comment{cite any number of Lande papers?}.  The latter does indeed require very weak selection.

@@

Almost uniquely among the stable laws, the Cauchy distribution has a probability density function expressible in elementary functions (the other exceptions being the L\'evy and Gaussian distributions): for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is 
\begin{equation}\label{eq:Cpdf}
	f(x;x_{0},\gamma) = \frac{1}{\pi\gamma\left[1+\left(\frac{x-x_{0}}{\gamma}\right)^{2}\right]}.  
\end{equation}	
Armed with this density function, we can explicitly compute the right hand size of \eqref{eqn:MSB_CF}, assuming that 
\begin{enumerate}[(i)]
\item the population is in mutation-selection balance corresponding to a Cauchy density with median 0 and scale $\gamma > 0$,
\item selection acts as $\mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for some $\delta > \gamma > 0$, and
\item the transmission noise $R$ is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(0,0,\delta-\gamma,0)$ law).
\end{enumerate}

Under these assumptions,  
\begin{align*}
	\frac{\mathbb{E}\left[e^{i \theta Z}(1+\mu(Z))\right]}{\mathbb{E}[1+\mu(Z)]} 
	&= e^{-\delta |\theta|}\\
	&= e^{-2\gamma \left|\frac{\theta}{2}\right|-(\delta-\gamma)|\theta|},
\end{align*} 
which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$, where $\phi(\theta) = e^{i\gamma|\theta|}$ and $\Phi(\theta) = e^{i(\delta - \gamma)|\theta|}$ are the characteristic functions of the population density and the transmission noise, respectively, giving us a.non-Gaussian solution to \eqref{eqn:MSB_CF} with both noise and selection.

>>>>

Unfortunately, once selection is included, \eqref{eqn:MSB_CF} is less amenable to the sort of simple arguments above.  We will content ourselves with presenting two examples of stabilizing selection, one already familiar to quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.

\paragraph{Gaussian Solution:}
First, consider the case when $\mu(x) = e^{a x^{2}}$.   In this case there is a Gaussian solution: suppose $Z$ is Gaussian with mean $m$ and variance $\sigma^{2}$.  Then,
$ \phi(\theta) = \exp(i m \theta - \frac{1}{2} \sigma^{2} \theta^{2}).  $
Moreover, provided $a < \frac{1}{2\sigma^{2}}$, we can explicitly compute $\frac{\mathbb{E}\left[e^{i \theta Z}\mu(Z)\right]}{\mathbb{E}[\mu(Z)]}$  using the Gaussian probability density function to conclude that
if $R$ is Gaussian with mean zero and variance $\eta^2$, then equation~\eqref{eqn:MSB_CF} becomes
\[
	e^{i m \theta - \frac{1}{2} \left(\frac{\sigma^{2}}{2} + \eta^{2}\right)\theta^{2}}
	= e^{i \frac{m}{1-2 a \sigma^{2}}\theta 
		- \frac{1}{2} \frac{\sigma^{2}}{1-2 a \sigma^{2}}\theta^{2}},
\]
which, equating real and complex components, has a solution provided $m = 0$ and
\[
	\sigma^{2} = \frac{\sqrt{16 a^{2} \eta^{4} + 24 a \eta^{2}+1} - 4a \eta^{2} -1}{4a} 
	=  2\eta^{2}(1-4a\eta^{2}) + O(a^{2}).
\]

Note that we do not require very weak selection to guarantee a Gaussian steady state, only that $a < \frac{1}{2\sigma^{2}}$.  However, this does not imply that the transient distribution is also Gaussian, as is sometimes assumed \citep[e.g.,][]{lande1976natural}. %  The latter does indeed require very weak selection.

@@

Almost uniquely among the stable laws, the Cauchy distribution has a probability density function expressible in elementary functions: 
% (the other exceptions being the L\'evy and Gaussian distributions): 
for a Cauchy random variable with median (location) $x_{0}$ and mean absolute deviation (scale) $\gamma$, the density is 
\begin{equation}\label{eq:Cpdf}
	f(x;x_{0},\gamma) = \frac{1}{\pi\gamma\left[1+\left(\frac{x-x_{0}}{\gamma}\right)^{2}\right]}.  
\end{equation}	
Armed with this density function, we can explicitly compute the right hand size of \eqref{eqn:MSB_CF}, assuming that 
\begin{enumerate}[(i)]
\item the population is in mutation-selection balance corresponding to a Cauchy density with median 0 and scale $\gamma > 0$,
\item selection acts as $\mu(x) = \frac{\delta^{2}}{\gamma^{2}} \frac{\gamma^{2} + x^{2}}{\delta^{2} + x^{2}}$ for some $\delta > \gamma > 0$, and
\item the Mendelian sampling term $R$ is given by a $\text{Cauchy}(0,\delta-\gamma)$ random variable (\ie corresponding to a stable $(1,0,\delta-\gamma,0)$ law).
\end{enumerate}

Under these assumptions, \eqref{eqn:MSB_CF} is
\begin{align*}
	\frac{\mathbb{E}\left[e^{i \theta Z}\mu(Z)\right]}{\mathbb{E}[\mu(Z)]} 
	&= e^{-\delta |\theta|}\\
	&= e^{-2\gamma \left|\frac{\theta}{2}\right|-(\delta-\gamma)|\theta|},
\end{align*} 
which we recognize as being equal to $\phi\left(\frac{\theta}{2}\right)^{2} \Phi(\theta)$, where $\phi(\theta) = e^{i\gamma|\theta|}$ and $\Phi(\theta) = e^{i(\delta - \gamma)|\theta|}$ are the characteristic functions of the population density and the Mendelian sampling term, respectively,
giving a.non-Gaussian solution to \eqref{eqn:MSB_CF} with both noise and selection.
